Military mechanization as a complex system

### 2018-07-15 paper2/paper2.tex
onward and upward

	Military mechanization is the degree to which a state's army is comprised of
	ground combat vehicles as opposed to infantry. Mechanization is both a 
	theoretically important topic in international relations scholarship and a practical 
	concern for policymakers. \citet{sechser2010army} influentially argue that
	inter-state dynamics (e.g., conflict) better explain mechanization choices than do
	intra-state factors (e.g., culture).

@@

This paper presents an agent-based model of military mechanization. It
reviews the academic literature on military mechanization. It discusses why
orthodox modeling strategies fail to fully describe international security competition
as a complex system and why agent-based modeling is needed. In support of these
claims, it ``agentizes'' the leading model of military mechanization and shows
that it does not hold. Finally, it presents a
novel agent-based model that better represents military mechanization.

@@

There are many potential causes of mechanization. Inter-state
competition and war is a prominent candidate. States may choose choose mechanization policies 
that they believe will allow them to prevail against potential adversaries, or to send a
signal of deterrence, or to learn from the perceived mistakes of the last war
\citep{mearsheimer1983conventional,huth1988extended,murray2011military}. 
Economic endowments may play a role: e.g., capital-rich states may mechanize more
than labor-rich states \citep{gartzke2001democracy}. Domestic political
institutions may matter: e.g., autocracies may be more willing than democracies to adopt
low-mechanization ``human wave'' tactics
\citep{quinlivan1999coup,reiter2002democracies,brooks2008shaping,talmadge2015dictator}. 
Ideology \citep{van1984cult} and national culture \citep[e.g.][]{pollack2004arabs} are
hard-to-measure but perhaps influential. Lastly,
organizational culture can play a role \citep{builder1989masks}. Militaries may pursue a
certain mechanization posture because it aligns with their institutional
objectives, which may or may not be coincident with rational strategy. 
Consider Ireland: between the 1920s and 1940s, it seriously considered expending a 
large portion of its defense budget on building a relatively small number of tanks. 
They would have been wholly insufficient to repel a British invasion (their ostensible purpose), 
and Ireland would have been better-served to pursue a low-mechanization guerrilla defense strategy, 
but tanks were seen as a prestigious signifier of a ``real'' professional
military and hence desirable
\citep{farrell1998professionalization,farrell2001transnational}.

\citet{sechser2010army} is the landmark study on military mechanization. They
assemble a longitudinal dataset at the country-year level of the mechanization
of all states between 1979 and 2001. They conduct regression analysis on this
panel data. They use a ordinary least squares (OLS) linear regression model with corrected standard
errors,

@@

with $y_{ij}$ the mechanization level of state $i$ in year $j$ (measured as the 
ratio of ground combat vehicles to infantry in a state's army) and $x_1 \ldots
x_j \ldots x_{P}$ covariates hypothesized to be associated with
mechanization. They conclude that, out of the
aforementioned potential causes, the first mentioned is most important: ``choices about
mechanization are strongly associated with a state's security environment.''

\section{System complexity and agent-based modeling}

Although \citet{sechser2010army} is a valuable contribution to the literature,
its orthodox multivariate statistical modeling is insufficient. Mechanization is a complex system. 
Agent-based modeling is a more appropriate paradigm for studying it. These three claims are addressed in turn.

@@

\citet{sechser2010army} operate on observational rather than experimental
data. The assignment of treatments to subjects is not randomized, causing
\textit{selection bias}. For example, \citet{sechser2010army} operationalize
dyadic ``rivalry'' as whether two states recently engaged in a Militarized
Interstate Dispute (MID) \citep{cow_midb}, and find significant association with mechanization. 
But interstate conflict is not a randomly assigned treatment. Rather, at least one state must select
themselves into it \citep{fearon2002selection}. States differ in their
proclivity for conflict initiation based on many distinct factors. 
Thus, correlation attributed to rivalry is in fact
bundling together many uncontrolled-for covariates, casting doubt on the claimed
finding. Natural experiments---a situation in which assignation of treatment is
known to be random---are the typical econometric strategy to avoid bias. For
example, \citet{angrist2002vouchers} exploit the fact that private school
vouchers were distributed by lottery in Colombia to obtain unbiased estimates of
the effect of schooling type (private vs. public) on educational achievement.
But such situations are rare in international relations---e.g., random assignation of war is an
absurd proposition---and so
\citet{sechser2010army} lack such a control. Given rife selection effects and few
natural experiments, econometric modeling alone is an insufficient tool for studying mechanization.

@@

The additive linear model in (1) assumes that each covariate $x$ is separable.  
Interpretation of the corresponding coefficient $\beta$ requires an assumption of independence 
(i.e. a lack of multicollinearity). It also makes an implicit claim about the direction of
causation: that mechanization $y$ is a function of those 
$x$.\footnote{To be clear,
\citet{sechser2010army} studiously avoid claiming causation in their text. But
their choice in what is modeled as $y$ and what is modeled $x$ says a lot about the perceived
direction of association.} These assumptions do not hold.

Let us take as an example GDP per capita. \citet{sechser2010army} find it negatively correlated with
mechanization at the $p < 0.05$ level. Unbiased estimation of this individual coefficient
requires lack of correlation between GDP per capita and other covariates, either individually or
jointly. Put differently, \textit{it requires that the system can be neatly decomposed into constituent parts}. Figure 1 is a
causal diagram depicting this logic. 
\begin{figure}[h]
	\centering
	\caption{Decomposable system}
	\begin{tikzpicture}
		\node[draw, rectangle](Mechanization)  at  (0,0) {Mechanization};	
		\node[draw, rectangle](GDP per capita)  at  (4,0) {GDP per capita};
		\node[draw, rectangle](Other covariates)  at  (4,-2) {Other covariates};
		\draw[->] (GDP per capita) -- (Mechanization);
		\draw[->] (Other covariates) -- (Mechanization);
	\end{tikzpicture}
\end{figure}

By standard statistical tests, the assumption of decomposability appears supported. The model's variance
inflation factor generally does not rise above 2, indicating a lack of
multicollinearity.\footnote{The high correlation between dummy variables}

However, the literature suggests that (1) is not a correctly specified model. GDP per capita is associated with
regime type (e.g., democracy vs. autocracy), and causality appears to flow in both directions, i.e. democracy
causes wealth and wealth also causes democracy \citep{helliwell1994empirical,barro1996democracy,robinson2006economic,doucouliagos2008democracy}. 
Regime type affects the probability of war: the closest thing to an empirical law in international relations is
that democracies do not fight each other
\citep{levy1988domestic}, while some kinds of autocracies are war-prone \citep{de1995war}. Regime type is widely argued to influence
battlefield performance in war because it affects the quality of military institutions and hence
capability of fielded forces \citep{reiter2002democracies,weeks2014dictators,talmadge2015dictator}. 
Battlefield performance affects regime type, e.g. a losing side may be subject to foreign-imposed
regime change a la 1939 France or 2003 Iraq
\citep{werner1996absolute,lo2008ensuring,downes2013forced}. \citet{sechser2010army} claim that the observed
historical probability of conflict (``rivalry'', in their terminology) affects mechanization; and claim that observed
battlefield performance (``lessons'') also affect mechanization. Mechanization surely affects the
probability of war, as it sends some kind of deterrent signal. Mechanization also surely affects
battlefield performance in war (if it did not, we would not care about it).
Mechanization also may affect GDP per capita in that states may over-spend on defense and 
impoverish the country, with the Soviet Union as a prominent example. Deterrence theory requires that 
battlefield performance affect the probability of war, i.e. militarily 
potent states are unlikely to be attacked. The exact truth value of these claimed relationships is
uncertain---in no small part due to the limitations of econometrics that were just addressed---but
they are generally uncontroversial statements in their respective fields.

@@

Figure 2 is the logic diagram resulting from the above. It must be emphasized that we are only scratching the 
surface of system complexity. For example, Figure 2 does not address the nature of relationships 
between covariates: linear or non-linear? subject to multi-way
interactions? But even as-is, Figure 2 visually conveys the essential point: \textit{the system is
not tractable to decomposition}. Ascertaining cause and effect in this system through simple observation is 
near-impossible. 
One instead must perform ``causal surgery'' \citep{pearl2009causality}, i.e. modify the value at one
of the nodes and observe the
resultant system state. But, as previously covered, doing so in international relations is an
absurd proposition: one cannot mete out international conflict in a real-world experimental
setting to assess the effect on mechanization. A modeling paradigm that would let us do so is needed.

@@

There is no law of nature dictating that an increase in one state's mechanization causes an adversary to 
raise its own mechanization. It also cannot be claimed that many individual choices aggregate into something 
akin to natural law---the sort of logic used 
in economics to justify treating supply and demand in a physics-esque manner---because the number of
people that decide mechanization choices is small. In a personalistic
dictatorship, it might be $n=1$; even in polities with more dispersed authority, such decisions
typically occur in the rarified and thinly populated realm of generals and cabinet ministers. 
\textit{A state's mechanization level changes because someone decides to change it}.

If one takes the previous statement to be true, then a model of military mechanization must be in
some sense a cognitive model. Some agent or set of agents perceives the outside world;

@@

Good for insurgency because its hidden, and

In the typology of Axtell \citep{axtell2000agents}

ABM is useful for exactly such settings.

Methodology studies
\citep{pepinsky2005agents,de2014agent,
gunitsky2013complexity,bousquet2011beyond,
harrison2012complexity,jervis1998system,modelski1990world,modelski1999long,cederman2001agent}

Insurgency studies
\citep{cioffi2010mason,weidmann2013violence,bhavnani2012modeling,bhavnani2014group}

Big IR \citep{cederman2010growing,earnest2008coordination,masad2016dissertation}

Fundamentals, e.g. cooperation \citep{axelrod1981evolution}

Just because \citep{epstein1996growing}

are ripe for agentization. The
conclusions drawn implicitly claim
that states are autonomous agents; that they have
cognition regarding how to perceive the outside world
and how to react to it; that spatiality and
neighborhood matter; that inter-agent relationships matter; and that
agents learn from experience. Agent-based modeling is uniquely well-suited to
represent such phenomena \citep{gilbert2005simulation,miller2009complex}.

@@

This paper is a novel contribution to the literature. Although agent-based
models have found traction in the study of grievance, insurgency, and
other population-centric security phenomena \citep[e.g.][]{epstein2002modeling},
they quite rarely are used to model international relations and state-centric security
phenomena despite some early work on this front
\citep[see][]{cederman1997emergent,axelrod1997complexity}. 
This paper's agentization of an orthodox security
studies model thus is novel in method. Its findings also are important. They
show that they show a solely ``external'', security
environment-focused view of mechanization does not stand up to agentization.
Spatiality, inter-agent relations, autonomous agents with cognition, and positive feedback
loops---all hallmarks of a complex system---drive results that significantly
diverge from theoretical expectations of the orthodox view.

@@

In summary, the orthodox view of mechanization is that it is a
reflexive process: states choose their mechanization posture largely based on
that neighbors, enemies, and recent military experiences (e.g. defeat in
counter-insurgency). This paper argues
that this view---based on traditional longitudinal data analysis---is ripe for
agentization. It implicitly argues for autonomous agents (states) that have
cognition regarding mechanization that takes into account other agents (e.g.
enemies), spatiality (e.g. neighbors), and learning from experience. This paper
builds such a model and validates it against the empirical record, using the
National Mechanization Index of \citet{sechser2010army}. It finds that the 
orthodox view of mechanization does not stand up to agentization. A positive feedback loop 
causes states to
mechanize much faster and to a greater degree than occurs in the real-world.
This implies the existence of some countervailing factor (possibly a negative
feedback loop) that mitigates the extent of mechanization. Candidates for what
this could be mostly center on ``internal'' factors---e.g., culture, economy,
and so on---as opposed to the ``external'', security environment-centric
orthodox view.
\end{document}

>>>>

	Military mechanization is the degree to which a state's army is comprised of
	ground combat vehicles as opposed to infantry. Mechanization is an important topic 
	in international relations scholarship and a practical concern for
	policymakers. This paper makes novel theoretical contributions by
	re-envisioning mechanization as a complex system. It also implements an
	agent-based model of military mechanization. The results of this model
	challenge the conventional wisdom that mechanization is primarily driven
	by external factors such as security environment. They imply that
	some internal factors (e.g. military culture or economic constraints)
	act as a ``brake'' on mechanization.

@@

This paper is divided into five sections. This \textbf{Introduction} gives
background on military mechanization. The \textbf{Theory} section rigorously
develops mechanization as a complex system best represented by
agent-based modeling. The \textbf{Model} section presents a hypothesis and
implements an agent-based model in Python to test that hypothesis.
\textbf{Results} are shown. Implications are explored in \textbf{Discussion}.

@@

There are many potential causes of mechanization. Inter-state
competition and war is a prominent candidate. States may choose choose mechanization policies 
that they believe will allow them to prevail against potential adversaries, or to send a
signal of deterrence, or to learn from the perceived mistakes of the last war
\citep{mearsheimer1983conventional,huth1988extended,murray2011military}. 
Economic endowments may play a role: e.g., capital-rich states may mechanize more
than labor-rich states \citep{gartzke2001democracy}. Domestic political
institutions may matter for military composition: e.g., autocracies may be more willing than democracies to adopt
low-mechanization ``human wave'' tactics
\citep{quinlivan1999coup,reiter2002democracies,brooks2008shaping,talmadge2015dictator}. 
Ideology \citep{van1984cult} and national culture \citep[e.g.][]{pollack2004arabs} are
hard-to-measure but perhaps influential. Lastly,
organizational culture can play a role \citep{builder1989masks}. Militaries may pursue a
certain mechanization posture because it aligns with their institutional
objectives, which may or may not be coincident with rational strategy.

\citet{sechser2010army} is the landmark study on military mechanization. They
assemble a longitudinal dataset at the country-year level of the mechanization
of most states between 1979 and 2001. They conduct regression analysis on this
panel data. They use a ordinary least squares (OLS) linear regression model with corrected standard
errors,

@@

with $y_{ij}$ the mechanization level of state $i$ in year $j$ (measured as the 
ratio of ground combat vehicles to infantry in a state's army) and $x_1 \ldots
x_j \ldots x_{P}$ covariates hypothesized to be associated with
mechanization. They conclude that, out of the
aforementioned potential causes, ``choices about
mechanization are strongly associated with a state's security environment.''
Their dataset and conclusions have guided further scholarship on military
mechanization, e.g. \citet{pilster2012democracies}, which mostly follow the same
statistical approach.

\section{Theory}

Definition of complex system. 
The orthodox multivariate statistical modeling approach is insufficient because
mechanization is a complex system. Agent-based modeling (ABM) is a more 
appropriate paradigm for studying it. These three
claims---insufficiency, complexity, and ABM---are addressed in turn.

@@

\citet{sechser2010army} use observational rather than experimental
data. The assignment of treatments to subjects is not randomized, causing
\textit{selection bias}. For example, they operationalize
dyadic ``rivalry'' as whether two states recently engaged in a Militarized
Interstate Dispute (MID), and find significant association with mechanization. 
But interstate conflict is not a randomly assigned treatment. Rather, at least one state must select
itself into it \citep{fearon2002selection}. States differ in their
proclivity for conflict, based on many factors. 
Thus, ``rivalry'' is in fact
bundling together many uncontrolled covariates, casting doubt on the claimed
finding. Similar problems arise for other variables included in the model.
Natural experiments---a situation in which treatment is
known to be randomly assigned---are the typical econometric strategy to avoid selection bias. For
example, \citet{angrist2002vouchers} exploit the fact that private school
vouchers were distributed by lottery in Colombia to obtain unbiased estimates of
the effect of schooling type (private vs. public) on educational achievement.
But such situations are rare in IR: e.g., random assignment of war is an
absurd proposition. \citet{sechser2010army} lack a good control
because it probably does not exist. Reliance on selection-riddled
observational data means that econometrics are at best an incomplete
tool for studying mechanization.

@@

The additive linear model in (1) assumes that each covariate $x$ is separable. Take as an example 
GDP per capita.\footnote{The example works as well with any
other covariate.} \citet{sechser2010army} find it negatively related with mechanization at the $p < 0.05$ level. 
But a valid interpretation requires an assumption
of its independence from other covariates (i.e. no multicollinearity).
It also makes an implicit claim about causal direction: that
mechanization $y$ is a function of various
$x$, and not the other way around.\footnote{To be clear,
\citet{sechser2010army} do not claim causation in their text. But
their choice in what is modeled as $y$ and what is modeled $x$ says a lot about the perceived
direction of association.} Put differently, \textit{it requires that the
observed system can be neatly 
decomposed into constituent parts}. Figure 1 is a causal diagram depicting this logic. 
\begin{figure}[h!]
	\centering
	\caption{Decomposable system}
	\begin{tikzpicture}
		\node[draw, rectangle](Mechanization)  at  (0,0) {Mechanization};	
		\node[draw, rectangle](GDP per capita)  at  (4,0) {GDP per capita};
		\node[draw, rectangle](Other covariates)  at  (4,-2) {Other covariates};
		\draw[->] (GDP per capita) -- (Mechanization);
		\draw[->] (Other covariates) -- (Mechanization);
	\end{tikzpicture}
\end{figure}

Initial statistical testing seems to support this assumption. The model's variance
inflation factor (VIF) does not indicate multicollinearity.\footnote{See the Appendix for
a supporting table. Some coefficients have VIF $> 8$, which strongly
indicates multicollinearity. However, this only occurs with ``dummy'' variables
and the variable to which they relate, and so is not problematic.} However, VIF
is contingent on the specific model being tested, and evidence suggests that 
(1) is not a correctly specified model. It over-simplifies or omits key
relationships.

A large literature suggests that GDP per capita is associated with
regime type (e.g., democracy vs. autocracy), and causality appears to flow in
both directions. Democracy causes wealth and wealth also causes democracy \citep{helliwell1994empirical,barro1996democracy,robinson2006economic,doucouliagos2008democracy}. 
Regime type affects the probability of war: the closest thing to an empirical law in international relations is
that democracies do not fight each other
\citep{levy1988domestic}, while some kinds of autocracies are war-prone \citep{de1995war}. 
Regime type is widely argued to influence
battlefield performance in war because it affects the quality of military institutions and hence
capability of fielded forces \citep{reiter2002democracies,weeks2014dictators,talmadge2015dictator}. 
Battlefield performance affects regime type, e.g. a losing side may be subject to foreign-imposed
regime change a la 1939 France or 2003 Iraq
\citep{werner1996absolute,lo2008ensuring,downes2013forced}. \citet{sechser2010army} claim that the observed
historical probability of conflict (``rivalry'', in their terminology) affects mechanization; and claim that observed
battlefield performance (``lessons'') also affect mechanization. Mechanization surely affects
battlefield performance in war; if it did not, we would not care about it
\citep[see also][]{lyall2009rage}.
Mechanization may affect GDP per capita in that states can over-spend on defense and 
impoverish the country
\citep{cappelen1984military,ward1992sizing,abu2003government}. Deterrence theory requires that 
battlefield performance affect the probability of war in that militarily 
potent states are unlikely to be attacked. The exact truth value of these claimed relationships is
uncertain---in no small part due to the limitations of econometrics that were just addressed---but
they are generally uncontroversial statements in their respective fields.

@@

Figure 2 is the logic diagram resulting from the above. It must be emphasized that we are only scratching the 
surface of system complexity. For example, Figure 2 does not address the nature of relationships 
between covariates: linear or non-linear? subject to multi-way
interactions? necessity and sufficiency distinctions? But even as-is, Figure 2 visually conveys the essential point: \textit{the system is
not tractable to decomposition}. Ascertaining cause and effect in this system
through simple observation is impossible because of endogeneity. Moreover, controlling for
confounding variables breaks down when the controlled-for variable directly
relates to the outcome of interest \citep[``post-treatment bias'',
c.f.][]{king2010hard}. One would like to perform ``causal surgery'' \citep{pearl2009causality}, i.e. modify the value at one
of the nodes and observe the resultant system state. But, as previously covered, one generally cannot conduct
experimental trials in international relations.  A modeling paradigm that would let us do so is needed.

@@

There is no law of nature dictating that an increase in one state's mechanization causes an adversary to 
raise its own mechanization. It also is not the case that many individual choices aggregate into something 
akin to natural law---the logic used 
in economics to justify treating supply and demand in a physics-esque manner---because the number of
people that decide mechanization choices is small. In a personalistic
dictatorship, it might be $n=1$. Even in polities with more dispersed authority, such decisions
typically occur in the rarified and thinly populated realm of generals and cabinet ministers. 
\textit{A state's mechanization level changes because someone decides to change it}.

If one takes the previous statement to be true, then a model of military
mechanization must be at least partly a cognitive model. Some agent or set of agents perceives the outside
world; processes that information; and takes action. Orthodox studies of
mechanization incorporate cognition to some degree. For example,
\citet{sechser2010army} test for learning: whether states that fail in
counter-insurgency wars decrease their mechanization as a result of this
``anti-mechanization lesson.'' This depiction, however, fails to capture 
key phenomena.

Human beings learn not only from their own experiences, but also from observing others.
State decisionmakers are no exception. For example, the stunning success of
Coalition precision-guided munitions in the 1991 Gulf War led other states who
were not direct combatants to reform their own militaries
\citep{scobell2011chinese}. There is no guarantee that such emulation always
works out. Consider Ireland: between the 1920s and 1940s, it sought to spend a
large portion of its defense budget on building a small number of
tanks, which at that time were novel technology being implemented by the leading
European militaries.
They would have been wholly insufficient to repel a British invasion (their ostensible purpose), 
and Ireland would have been better-served to pursue a low-mechanization guerrilla defense strategy, 
but tanks were seen as a prestigious signifier of a ``real'' professional
military and hence desirable
\citep{farrell1998professionalization,farrell2001transnational}. Such learning
is absent from orthodox models of military mechanization.

@@

In light of the above, agent-based modeling (ABM) is 
uniquely well-suited to study mechanization. Computer simulations provide the possibility 
to conduct controlled experiments \textit{in silico} \citep{epstein1996growing} that are not possible 
in the real world, freeing us from the tyranny of selection bias. Because the
modeler can adjust various ``dials'' and observe the outcome, ABM allows causal
inference or at least insight even in highly complex, perhaps unsolvable systems
\citep[c.f.][]{axtell2000agents}. ABM focuses on autonomous agents with
cognition, for which even relatively simple behavioral rules can generate
behavior at least as complex as mechanization
\citep{gilbert2005simulation,miller2009complex}.

Complex systems and ABM have been used in IR. Many methodology-focused
works have identified their utility to the field \citep[some prominent examples
are][]{modelski1990world,jervis1998system,pepinsky2005agents,de2014agent}.
Although actual usage has somewhat lagged behind theoretical enthusiasm,
ABM has been used to explore emergent 
balance of power \citep{cederman1997emergent}, international negotiation
\citep{earnest2008coordination}, and international crises and conflict
\citep{masad2016dissertation}. It is especially popular for studying
population-centric security phenomena such as terrorism and insurgency
\citep{cioffi2010mason,bhavnani2012modeling,weidmann2013violence,bhavnani2014group}.

The author's sense is that adoption of ABM in IR has lagged because it is
regarded as non-empirical. ABM offers no guarantees: just
because one's model ``works'' (internally consistent, externally plausible) 
does not mean it is true. Econometrics are the dominant methodological approach 
in quantitative IR because they seem to offer an objective yardstick for competing hypotheses: 
does A or B best fit the data? This paper contends that
epistemological limitations (most importantly, selection bias stemming from
non-randomization, and system complexity) mean that econometrics do not
provide reliable statistical inference for certain classes of IR problems, among
them mechanization \citep[c.f.][for a concurring opinion]{king2010hard}. 
If neither econometrics nor ABM can offer 
certitude that a hypothesis is true---as is the case for mechanization---then the latter 
should be preferred for its advantages in experimentation, transparency, researcher control, 
and amenability to complex systems.

@@

This paper is a novel contribution to the literature. Although agent-based
models have found traction in the study of grievance, insurgency, and
other population-centric security phenomena \citep[e.g.][]{epstein2002modeling},
they quite rarely are used to model international relations and state-centric security
phenomena despite some early work on this front
\citep[see][]{,axelrod1997complexity}. 
This paper's agentization of an orthodox security
studies model thus is novel in method. Its findings also are important. They
show that they show a solely ``external'', security
environment-focused view of mechanization does not stand up to agentization.
Spatiality, inter-agent relations, autonomous agents with cognition, and positive feedback
loops---all hallmarks of a complex system---drive results that significantly
diverge from theoretical expectations of the orthodox view.

@@

In summary, the orthodox view of mechanization is that it is a
reflexive process: states choose their mechanization posture largely based on
that neighbors, enemies, and recent military experiences (e.g. defeat in
counter-insurgency). This paper argues
that this view---based on traditional longitudinal data analysis---is ripe for
agentization. It implicitly argues for autonomous agents (states) that have
cognition regarding mechanization that takes into account other agents (e.g.
enemies), spatiality (e.g. neighbors), and learning from experience. This paper
builds such a model and validates it against the empirical record, using the
National Mechanization Index of \citet{sechser2010army}. It finds that the 
orthodox view of mechanization does not stand up to agentization. A positive feedback loop 
causes states to
mechanize much faster and to a greater degree than occurs in the real-world.
This implies the existence of some countervailing factor (possibly a negative
feedback loop) that mitigates the extent of mechanization. Candidates for what
this could be mostly center on ``internal'' factors---e.g., culture, economy,
and so on---as opposed to the ``external'', security environment-centric
orthodox view.

are ripe for agentization. The
conclusions drawn implicitly claim
that states are autonomous agents; that they have
cognition regarding how to perceive the outside world
and how to react to it; that spatiality and
neighborhood matter; that inter-agent relationships matter; and that
agents learn from experience. Agent-based modeling is uniquely well-suited to
represent such phenomena .

\begin{table}[h]
	\centering
	\begin{tabular}{l r}
		\hline
		\textbf{Variable} & \textbf{VIF} \\ 
		\hline
		log\_mech\_enemies & 8.76 \\ 
		enemies\_dummy & 9.12 \\ 
		log\_mech\_neighbors & 2.19 \\ 
		neighbors\_dummy & 2.03 \\ 
		log\_mountainous & 1.26 \\ 
		antimech\_lesson\_2 & 1.21 \\ 
		promech\_lesson\_2 & 1.11 \\ 
		democracy & 1.69 \\ 
		military\_govt & 1.13 \\ 
		instability & 1.12 \\ 
		log\_gdp\_cap & 2.10 \\ 
		iron\_steel & 1.32 \\ 
		log\_mech\_allies & 12.64 \\ 
		allies\_dummy & 12.59 \\ 
		cold\_war\_bloc & 1.48 \\ 
		igos & 2.03 \\ 
		post\_soviet & 1.33 \\ 
		log\_mech\_lag2 & 1.57 \\ 
		\hline
	\end{tabular}
\end{table}
\end{document}
